# Reproducible Research: Peer Assessment 1

# Loading required packages
```{r Load Packages, echo=TRUE, results='hide', message=FALSE}
require(ggplot2)
```

## Loading and preprocessing the data
```{r Load Data,echo=TRUE, cache=TRUE}
setwd("/Users/sharannarang/Desktop/R projects/Reproducible Research/RepData_PeerAssessment1/")
unzip(zipfile="activity.zip")
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
In order to compute the total number of steps per day, I have created a new data frame (activity.na) and replaced the NA values with 0. 
```{r Total Steps/day, echo=TRUE, cache=FALSE, message=FALSE}
activity.na <- activity
activity.na[is.na(activity)] <- 0
total.steps <- tapply(activity.na$steps, activity.na$date, sum)
qplot(total.steps, main="Total Steps/day", xlab="Total steps") + scale_y_discrete(breaks=seq(0,12,by=2))
```

```{r Mean Total/Day}
mean.total <- mean(total.steps)
median.total <- median(total.steps)
```

The average total steps taken per day is `r mean.total`. The median total steps per day is `r median.total`.

## What is the average daily activity pattern?


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?